---
title: "Project 3: stat302Project tutorial"
author: "Lauren Ng"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stat302Project tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Installation
To download the corncob package, use the code below.

```{r eval=FALSE}
install.packages("stat302Package")
library(stat302Package)
```

Alternatively, you can install the development version directly from GitHub.

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("laurenng/stat302Package")
library(stat302Package)
```

# Use

The vignette demonstrates example usage of all main functions. Please file an issue if you have a request for a tutorial that is not currently included. You can see the vignette by using the following code:

```{r eval=FALSE}
library(stat302Project)
# Use this to view the vignette in the corncob HTML help
help(package = "stat302Project", help_type = "html")
# Use this to view the vignette as an isolated HTML file
utils::browseVignettes(package = "stat302Project")
```

# Tutorial for my_t.test
```{r, include = FALSE}
# head(my_gapminder)

# lifeExpect <- my_gapminder$lifeExp


# two sided: 
# my_t.test(lifeExpect, "two-sided", 60)
```


# Tutorial for my_lm


# Tutorial for my_knn_cv


# Tutorial for my_rf_cv


```{r setup}
library(stat302Package)
```
